Classification Model of Exercise Movement Using Random Forests
========================================================
```{r}
train.raw = read.csv(file="train.csv", na.string=c("NA", "")
dim(train.raw)
```

### reduce the size to facilitate analysis and model creation time
```{r}
sampIndx = sample(nrow(train.raw), 4000)
train.small = train.raw[sampIndx,]
```

### clean the dataset of missing and NA values
```{r}
NArm = apply(train.small, 2, function(x) {sum(is.na(x))})
train.sm.valid = train.small[,which(NArm == 0)]
```

### remove noisy and nor information variables
```{r}
rmIndex <- grep("timestamp|X|user_name|new_window",names(train.sm.valid))
train.sm.valid = train.sm.valid[,-rmIndex]
```

# split dataset into train and validation sets
```{r}
inTrain = createDataPartition(y=train.sm.valid$classe, p=0.7, list=FALSE)
TRAIN = train.sm.valid[inTrain,]
VALIDATE = train.sm.valid[-inTrain,]
```

# exploratory, confirm the size of the datasets
```{r}
dim(TRAIN)
dim(VALIDATE)
```

# 4 level cross-validation to expedite model construction
```{r}
trControl=trainControl(method="cv", number=4, allowParallel=TRUE)
```

# train model using random forest algorithm
```{r}
randForMod = train(classe ~ ., method="rf", data=TRAIN, trControl)

PREDS = predict(randForMod, VALIDATE)
```

# estimate out of sample accuracy using the validation set
```{r}
confusionMatrix(PREDS, VALIDATE$classe)
```




